Forecasting the United States 2024 Presidential Election

### 2024-11-04 paper/paper.qmd
updated paper to include appendix graphs

plot_path <- here::here("other", "plots", "election_map.pdf")

#read in cleaned data
cleaned_poll_data <-
  read_csv(file = here("data/cleaned_poll_data.csv"), 
           show_col_types = FALSE)
swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))
```

@@

Posterior predictive checks were conducted to assess the model's ability to replicate observed data. Density overlay plots for both Trump and Harris models indicated reasonable fits, although some discrepancies were noted near the peaks of the distributions. Specifically, the Trump model's replicated data tended to be slightly smaller around the peak, while the Harris model exhibited more extreme differences at the peak. Residual plots appeared randomly scattered, suggesting no major systematic biases, although some clumping was observed, indicating areas where the model may fit less effectively [REF].

@@

```{r}
#| fig-cap: Election Prediction Results
#| echo: false
#| warning: false
#| message: false

#| echo: false
#| warning: false
#| message: false
election_results <- data.frame(
  Metric = c("Chance of Winning", "Average Electoral Votes"),
  Harris = c("67.4%", "279.33"),
  Trump = c("31.7%", "258.67"),
  stringsAsFactors = FALSE
)

@@

# Display a nicely formatted table
kable(election_results, caption = "Election Prediction Results")

@@

Based on the simulated elections, this map illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.[Appendix]

```{r}
#| fig-cap: Election Map
#| echo: false
#| warning: false
#| message: false
knitr::include_graphics(plot_path)

```
The results of the US elections in large part hinge on the direction of the swing states. A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris. [FIG] shows the probability that both candidates have of winning each of the 7 swing states.

```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
#| echo: false
#| warning: false
#| message: false
# Reshape the data for plotting
swing_long <- swing_probs %>%
  pivot_longer(cols = c("Trump", "Harris"), names_to = "Candidate", values_to = "Probability")

@@

The wording of questions should be straightforward and easy to understand, minimizing error in measurement caused by misinterpretation of questions. The question should also be asked neutrally, without leading participants to any particular response. Likert scales, including an option for “don’t know”, should be used for all questions for ease and consistency of participant understanding. As well, keeping questions and answer options simple minimizes the burden placed on participants, which reduces the attrition caused by participants losing interest in the poll [@citeTB]. The “don’t know” option is also effective in helping reduce the decision difficulty to certain participants who are truly unsure. Before officially deploying the survey with our selected sample, it first needs to be tested to catch ambiguous or other poorly phrased questions [@citeTB]. To encourage participation in this research, a financial incentive should be offered for full completion of the poll.

After analysis of poll responses, the votes of the historically consistent states should be added back into consideration before making the final forecast for election results.

@@

```
\newpage

>>>>

#read in cleaned data
cleaned_poll_data <-
  read_csv(file = here("data/cleaned_poll_data.csv"), 
           show_col_types = FALSE)
swing_probs <- read_csv(file = here("other/plots/swing_states_probabilities.csv"))

state_win_counts <- read_csv(file = here("other/plots/state_win_counts.csv"))
```

@@

Posterior predictive checks were conducted to assess the model's ability to replicate observed data[REF]. Density overlay plots for both Trump and Harris models indicated reasonable fits, although some discrepancies were noted near the peaks of the distributions. Specifically, the Trump model's replicated data tended to be slightly smaller around the peak, while the Harris model exhibited more extreme differences at the peak.

@@

```{r}
#| label: fig-4
#| fig-cap: Election Prediction Results
#| echo: false
#| warning: false
#| message: false

election_results <- data.frame(
  Metric = c("Chance of Winning", "Average Electoral Votes"),
  Harris = c("67.4%", "279.33"),
  Trump = c("31.7%", "258.67"),
  stringsAsFactors = FALSE
)

@@

@@

Based on the simulated elections, @fig-5 illustrates the most likely winner of each state. In this scenario Harris wins 276 Electoral Votes, and Trump wins 262.

```{r}
#| label: fig-5
#| fig-cap: Election Map
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","election_map.pdf" )
knitr::include_graphics(plot_path)

```
The results of the US elections in large part hinge on the direction of the swing states. A swing state is a state that holds a significant number of electoral votes and could reasonable be won by either Trump or Harris. @fig-6 shows the probability that both candidates have of winning each of the 7 swing states.

```{r plot-probabilities, echo=FALSE, fig.width=8, fig.height=6}
#| label: fig-6
#| echo: false
#| warning: false
#| message: false
# Reshape the data for plotting
swing_long <- swing_probs %>%
  pivot_longer(cols = c("Trump", "Harris"), names_to = "Candidate", values_to = "Probability")

@@

- questions (wording, order, options)
- treatment of non-response
- offer of incentive

After analysis of poll responses, the votes of the consistent states should be added back into consideration before making the final forecast for election results.

@@

## State By State Win {#sec-trace-plots}

```{r}
#| fig-cap: Number of Wins in 1000 Simulations by State
#| echo: false
#| warning: false
#| message: false

kable(state_win_counts, caption = "Election Results by State", align = 'lccc') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE)

## Trace Plots {#sec-trace-plots}

```{r}
#| fig-cap: Trump Trace Plots
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","trump_trace.pdf" )
knitr::include_graphics(plot_path)

```{r}
#| fig-cap: Harris Trace Plots
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","harris_trace.pdf" )
knitr::include_graphics(plot_path)

## Posterior Predictive Checks {#sec-ppc}

```{r}
#| fig-cap: Trump Posterior Predictive Checks
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","trump_ppc.pdf" )
knitr::include_graphics(plot_path)

```{r}
#| fig-cap: Harris Posterior Predictive Checks
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","harris_ppc.pdf" )
knitr::include_graphics(plot_path)

## Residual Plots {#sec-residual-plots}

```{r}
#| fig-cap: Trump Residuals
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","trump_residuals.pdf" )
knitr::include_graphics(plot_path)

```{r}
#| fig-cap: Harris Residuals
#| echo: false
#| warning: false
#| message: false

plot_path <- here::here("other", "plots","harris_residuals.pdf" )
knitr::include_graphics(plot_path)
